\documentclass[11pt, a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{braket}
\usepackage{graphicx}

\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

\newcommand{\Hgen}{\hat{\mathcal{H}}}
\newcommand{\Kry}{\mathcal{K}}
\newcommand{\Rset}{\mathbb{R}}
\newcommand{\Cset}{\mathbb{C}}
\newcommand{\Eblr}{\mathbb{E}}
\newcommand{\Kop}{\hat{K}}

\title{\textbf{Wavefunction Flows in the \texttt{unfer} Kernel:\\
An Analytical, Grid-Free Realization of the Continuity Hamiltonian
via Orthogonal-Fock Encoding and Inversion-Free Krylov Reduction}\\[4pt]
\large A Self-Contained, Pedagogical Account}
\author{Implementation Architecture Report}
\date{June 2026 (rev.\ 5, with Tomographic Subspace Recovery, rev 14 SIRK-backed unitary flow, rev 15 module demo + benchmarks, rev 16 Quantum Bayesian Update implementation, rev 17 full SIRK-whitened Krylov basis + kernel surface wiring + 5th module demo)}

\begin{document}

\maketitle

\begin{abstract}
Layden, Sweke, Havl\'i\v{c}ek, Chowdhury, and Neklyudov~\cite{Layden2025}
recently showed that classical \emph{flow models} map exactly onto the
Schr\"odinger equation: the qsample $\Psi_t(x)=\sqrt{p_t(x)}$ of a flow's
marginal density obeys $i\,\partial_t\Psi_t=\Hgen_t\Psi_t$ for an unusual
\emph{continuity Hamiltonian} $\Hgen_t=\tfrac12[\hat p\cdot v_t(\hat x)+
v_t(\hat x)\cdot\hat p]$, which for a conservative field $v_t=\nabla V_t$
takes the elegant commutator form $\Hgen_t^c=i[\Kop,V_t(\hat x)]$. Crucially,
this generator is \emph{fully specified by the already-learned classical
velocity field} --- no further training, parameterization, or variational
optimization. Their algorithm prepares qsamples on a \emph{digital quantum
computer} by simulating $\Hgen_t^c$ with Fourier collocation and product
formulas, enabling, e.g., quadratically faster mean estimation. This report
gives a self-contained, pedagogical account of how the \texttt{unfer}
probability kernel realizes the \emph{same} wavefunction flow
\emph{classically}, with no spatial grid. We replace the $N^d$ collocation
grid by an \emph{analytical orthogonal-Fock encoding}: each data point becomes
a mutually orthogonal single-particle mode, so the conservative potential
\emph{decouples} and its flow-matching parameters reduce to a diagonal,
closed-form $\mathcal{O}(M)$ fit for $M$ data points. The resulting bounded
generator is compressed by the Shift-Invert Rational Krylov (SIRK) method of
Hashimoto and Nodera~\cite{Hashimoto2019} --- replacing the paper's product
formulas --- giving exponential approximation-error decay and reducing
single-step inference to an $\mathcal{O}(m^2)$ matrix exponential with
$m\ll M$. We also make explicit the measure-theoretic foundation: in the
infinite-dimensional limit the uniform measure on a hypersphere becomes the
Gaussian measure, and the Gegenbauer polynomials that diagonalize it become
the Hermite polynomials (L\'opez and Temme, 1999; Mehler, 1866), which
identifies the Gaussian source $p_0$ with the Fock vacuum and any orthogonal
excitation with acquired information. This second revision (June 2026) adds
two new contributions. First, the Hermitian off-diagonal vacuum--data coupling
$\Hgen_{\mathrm{off}}=\ket{0}\bra{0}+\tfrac12\sum_j\bar\alpha_j(B_j^\dagger
P_0+P_0 B_j)$, which realizes coherent (Rabi-like) amplitude transport
between the source and the data channels in place of the stationary diagonal
surrogate, while preserving the Hermiticity required by the SIRK pipeline.
Second, the full \emph{Tomographic Subspace Recovery} (TSR) pipeline: a
two-level Count-Sketch hashing scheme ($S_1:\Rset^d\to\Rset^k$, $S_2$:
features $\to$ $K_2$-dim single-excitation Fock states) that decouples
semantics from reasoning, pre-projects all raw-coordinate observables into
the $m^2$ Krylov operator basis, and executes online generation in four
phases (encode, evolve, tomographic reconstruct, lossless decode) with
\emph{no} $\mathcal{O}(M)$ dependence at inference. We close by detailing
the integration into \texttt{unfer}'s GPU-accelerated \texttt{fock\textunderscore
sirk} solver, its \texttt{nested\textunderscore fock\textunderscore algebra}
CAS, the new \texttt{qfm} tomographic crate, and the Austral module runtime.
\end{abstract}

\tableofcontents

\section{Introduction}

\subsection{Flow models and the qsample bridge}

\emph{Flow models} are a cornerstone of modern generative machine learning.
They learn a continuous-time Markov process --- realized on an ordinary
(classical) computer --- that transports a simple source distribution $p_0$
(typically isotropic Gaussian noise) into a complex target $p_{\mathrm{targ}}$
from which we are given samples. The process is specified by a time-dependent
\emph{velocity field} $v_t(x)$ through the ordinary differential equation (ODE)
\begin{equation}
  \frac{\mathrm{d}}{\mathrm{d}t}\,x_t=v_t(x_t),\qquad x_0\sim p_0,
  \label{eq:flowode}
\end{equation}
whose random initial condition makes $x_t\sim p_t$ a random variable. The
family of marginals $(p_t)_{t\in[0,T]}$ is the \emph{probability path}, and by
the change-of-variables formula it satisfies the \emph{continuity equation}
\begin{equation}
  \partial_t\,p_t(x)=-\nabla\!\cdot\bigl[v_t(x)\,p_t(x)\bigr].
  \label{eq:continuityeq}
\end{equation}
Flow models learn $v_t^\theta\approx v_t$ (e.g.\ by flow matching,
Section~\ref{sec:flowprimer}) so that integrating~\eqref{eq:flowode} from
$x_0\sim p_0$ yields $x_T\sim p_T\approx p_{\mathrm{targ}}$.

In quantum computing, the corresponding object of interest is the
\emph{qsample}: the coherent encoding
\begin{equation}
  \ket{\Psi_p}=\sum_x\sqrt{p(x)}\,\ket{x}
  \quad\text{(discrete)},\qquad
  \ket{\Psi_p}=\int\sqrt{p(x)}\,\ket{x}\,\mathrm{d}x
  \quad\text{(continuous)}.
\end{equation}
Measuring $\ket{\Psi_p}$ in the $\{\ket{x}\}$ basis returns a sample
$x\sim p$, but \emph{preparing} a qsample is strictly more powerful than
sampling: with qsample access one can build quantum algorithms for mean
estimation and property testing with provable sample-complexity advantages
(Section~\ref{sec:qsample}). Preparing qsamples for arbitrary $p$ is hard in
general, so the central question is: \emph{for which $p$ can we prepare
qsamples efficiently?}

\subsection{Wavefunction flows (Layden et al.)}

Layden, Sweke, Havl\'i\v{c}ek, Chowdhury, and Neklyudov~\cite{Layden2025}
answer this for the vast family of distributions expressible by flow models.
Their key observation is a striking identity: if $p_t$ obeys the continuity
equation~\eqref{eq:continuityeq}, then its qsample wavefunction
$\Psi_t(x)=\sqrt{p_t(x)}$ obeys the Schr\"odinger equation
$i\,\partial_t\Psi_t=\Hgen_t\Psi_t$ for the \emph{continuity Hamiltonian}
\begin{equation}
  \Hgen_t=\tfrac12\bigl[\hat p\cdot v_t(\hat x)+v_t(\hat x)\cdot\hat p\bigr],
  \label{eq:continuityH}
\end{equation}
derived in Section~\ref{sec:continuity}. The dynamics it generates they call a
\emph{wavefunction flow}. Two features make this remarkable:
\begin{enumerate}
\item \textbf{No further learning.} $\Hgen_t$ is \emph{completely determined by
the already-learned classical velocity field} $v_t^\theta$. There is no
parameterized quantum circuit, no variational loop, no barren-plateau
pathology --- the entire ``training'' has been done classically, and the
quantum side is pure Hamiltonian simulation.
\item \textbf{Provable qsample preparation.} Evolving the easily prepared
initial qsample $\Psi_0=\sqrt{p_0}$ (a tensor product of standard-normal
qsamples for Gaussian $p_0$) under $\Hgen_t$ for time $T$ produces exactly
$\Psi_T=\sqrt{p_T}$; the distance to the ideal target $\sqrt{p_{\mathrm{targ}}}$
is controlled \emph{entirely classically} by $\|p_T-p_{\mathrm{targ}}\|$.
\end{enumerate}
The paper's main technical work is then turning this into a digital quantum
algorithm: discretizing space (Fourier collocation on an $N^d$ grid) and time
(product formulas), with error bounds (their Theorems~1 and~2).

\subsection{This work: a classical, grid-free realization in \texttt{unfer}}

The Layden et al.\ route targets a \emph{digital quantum computer} and pays for
it with an $N^d$ spatial collocation grid and a product-formula time
discretization whose cost grows with the operator norm $\|H_t^c\|=
\mathcal{O}(N^2 d/L^2)$. The \texttt{unfer} probability kernel already carries
all the machinery to realize the \emph{same} continuity-Hamiltonian
wavefunction flow \emph{classically}, on a separable Fock space, with Born-rule
semantics. This report shows how, and in doing so removes the spatial grid
altogether. The construction rests on three pillars, developed pedagogically
in Sections~\ref{sec:continuity}--\ref{sec:krylov}:

\begin{enumerate}
\item \textbf{Orthogonal Fock encoding (Section~\ref{sec:framework}).} Instead
of an $N^d$ collocation grid, encode each of the $M$ data points as a distinct,
mutually orthogonal single-boson mode. For a conservative field the potential
$V_t$ then becomes \emph{diagonal}, the flow-matching normal equations decouple
into $M$ independent scalar equations, and ``training'' is an $\mathcal{O}(M)$
closed-form average with zero data loss --- grid-free and inversion-free.

\item \textbf{Inversion-free Krylov reduction (Section~\ref{sec:krylov}).} In
place of the paper's product-formula time stepping, compress the resulting
time-independent generator $\bar H$ with the Shift-Invert Rational Krylov
(SIRK) method of Hashimoto and Nodera~\cite{Hashimoto2019}, whose error decays
\emph{exponentially} in the subspace dimension $m$. Because the potential has
no cross-terms, each Krylov step costs $\mathcal{O}(M)$ and inference is a
single $\mathcal{O}(m^2)$ matrix exponential.

\item \textbf{The Gaussian-source $=$ Fock-vacuum identification
(Section~\ref{sec:mehler}).} The Mehler/L\'opez--Temme limit of Gegenbauer to
Hermite polynomials makes precise why the Gaussian source $p_0$ of a flow model
is exactly the Fock vacuum, and why orthogonal excitations carry information.
\end{enumerate}

Section~\ref{sec:impl} maps every piece onto the existing \texttt{unfer}
modules. Throughout we keep prerequisites minimal: a reader comfortable with
ladder operators, Gaussian integrals, and the matrix exponential should be able
to follow without external references. We are explicit about scope: the shipped
\texttt{unfer} builtin implements a spectrally-stable diagonal surrogate of the
full off-diagonal continuity generator, and we flag the difference precisely
(Sections~\ref{sec:framework} and~\ref{sec:impl}).

\section{Background I: flow matching and the qsample}
\label{sec:flowprimer}

\subsection{Learning the velocity field by flow matching}

A flow model needs a velocity field $v_t$ transporting $p_0$ to
$p_{\mathrm{targ}}$. The ground-truth $v_t$ that generates a chosen probability
path is unknown, so the naive flow-matching loss is intractable. The decisive
trick~\cite{Layden2025} is the \emph{conditional flow matching} (CFM) loss,
which has the same gradient but is a tractable regression. For the
straight-line (optimal-transport) interpolation
$x_t=(1-t)x_0+t\,x_{\mathrm{targ}}$ with constant conditional velocity
$x_{\mathrm{targ}}-x_0$,
\begin{equation}
  \mathcal{L}_{\mathrm{CFM}}(\theta)=
  \Eblr_{t\sim\mathrm{unif}[0,1]}\,
  \Eblr_{x_0\sim p_0}\,
  \Eblr_{x_{\mathrm{targ}}\sim p_{\mathrm{targ}}}
  \Bigl\|v_t^\theta\bigl[(1-t)x_0+t\,x_{\mathrm{targ}}\bigr]
  -\bigl(x_{\mathrm{targ}}-x_0\bigr)\Bigr\|^2 .
  \label{eq:cfm}
\end{equation}
This estimates the gradient of the true flow-matching loss by an unbiased Monte
Carlo average over training samples and \emph{never} solves a differential
equation. The wavefunction-flow construction is agnostic to \emph{how} $v_t$ is
learned; it only requires that some $v_t^\theta\approx v_t$ exists. In
Section~\ref{sec:framework} we exploit~\eqref{eq:cfm} directly: under the
orthogonal-Fock encoding its minimizer is available in closed form.

\subsection{Diffusion models are flow models}

Diffusion models~\cite{Layden2025} provide another route to $v_t^\theta$. A
forward stochastic differential equation $\mathrm{d}x_t=f_t(x_t)\,\mathrm{d}t+
g_t\,\mathrm{d}w_t$ drives $p_{\mathrm{targ}}$ to a Gaussian; the associated
\emph{probability-flow ODE} replaces the reverse SDE by a deterministic field
\begin{equation}
  v_t(x)=f_t(x)-\tfrac12 g_t^2\,\nabla\ln p_t(x),
\end{equation}
which is an instance of~\eqref{eq:flowode}. When $f_t$ is conservative so is
$v_t$, so diffusion models fall under the same conservative-field analysis we
use below. This is why the single construction covers both of the dominant
continuous-variable generative paradigms.

\subsection{Qsamples and their statistical payoff}
\label{sec:qsample}

The reason to lift $p$ to its qsample $\Psi=\sqrt p$ is statistical leverage.
For mean estimation of $f(x)$ under $x\sim p$, the best classical estimators
(sub-Gaussian, e.g.\ median-of-means) achieve, with $m$ samples and failure
probability $\Delta$,
\begin{equation}
  \Pr\Bigl[|\tilde\mu-\mu|>C\sigma\sqrt{\tfrac{\log(1/\Delta)}{m}}\Bigr]\le\Delta,
  \qquad \sigma=\sqrt{\mathrm{Var}_{x\sim p}f(x)} .
\end{equation}
Given qsample-preparation access, the quantum sub-Gaussian
estimator~\cite{Layden2025} instead achieves
\begin{equation}
  \Pr\Bigl[|\tilde\mu-\mu|>\sigma\,\tfrac{\log(1/\Delta)}{m}\Bigr]\le\Delta,
\end{equation}
a \emph{quadratic} speed-up in the dependence on $m$, with no promise on the
variance beyond finiteness. Wavefunction flows make exactly this access
available for any distribution expressible by a (conservative) flow model. The
\texttt{unfer} kernel exposes the analogous probabilities through its Born-rule
layer, so the same statistical quantities are computable on the classical
substrate.

\section{Background II: operator functions and the SIRK method}

Realizing a wavefunction flow --- on a quantum computer or on \texttt{unfer} ---
ultimately asks us to apply $e^{-i\Hgen t}$ (or, more generally, an operator
$\varphi$-function) to a state. The paper~\cite{Layden2025} does this with
product formulas; \texttt{unfer} does it with the Shift-Invert Rational Krylov
(SIRK) method of Hashimoto and Nodera~\cite{Hashimoto2019}, which we summarize
here because it is the engine of Section~\ref{sec:krylov}.

\subsection{Exponential integrators and \texorpdfstring{$\varphi$}{phi}-functions}

Evolution equations $u'=Au+g$ have the variation-of-constants solution
\begin{equation}
  u(t)=e^{tA}u_0+\int_0^t e^{(t-s)A}g(s)\,\mathrm{d}s .
\end{equation}
\emph{Exponential integrators} discretize the integral using the family of
$\varphi$-functions
\begin{equation}
  \varphi_0(z)=e^z,\qquad
  \varphi_k(z)=\int_0^1 e^{sz}\,\frac{(1-s)^{k-1}}{(k-1)!}\,\mathrm{d}s
  \quad(k\ge 1),
\end{equation}
e.g.\ the one-step scheme $u_i=e^{\delta A}u_{i-1}+\delta\,\varphi_1(\delta
A)\,g_{i-1}$. The computational kernel is therefore the product
$\varphi_k(A)\,v$ of an operator function with a vector --- exactly the
primitive that wavefunction-flow inference needs (with $k=0$, $A=-i\Hgen$).

\subsection{Why naive (Arnoldi) Krylov fails for stiff generators}

The Arnoldi method builds the polynomial Krylov subspace
\begin{equation}
  \Kry_m(A,v)=\mathrm{span}\{v,Av,\dots,A^{m-1}v\},
\end{equation}
and approximates $\varphi_k(A)v\approx V_m\,\varphi_k(\mathbf{H}_m)\,V_m^*v$,
where $V_m$ holds an orthonormal basis and
$\mathbf{H}_m=V_m^*AV_m$ is the compression of $A$. For a \emph{bounded} $A$
the error decays geometrically, governed by polynomial approximation of
$\varphi_k$ on the numerical range $W(A)$. But the continuity Hamiltonian
contains the kinetic operator $\Kop=-\tfrac12\nabla^2$, which is
\emph{unbounded}: $W(A)$ is unbounded, the relevant conformal map ceases to
exist, and the error is no longer guaranteed to decay in $m$. This is the
stiffness obstruction --- the same one the paper~\cite{Layden2025} confronts
when it notes $\|K\|=\mathcal{O}(N^2 d/L^2)$ blows up with grid resolution.

\subsection{Resolvents, rational Krylov, and SIRK}

The cure is to work with the \emph{resolvent} $X=(\gamma I-A)^{-1}$, which is
bounded whenever $\gamma$ is outside the spectrum of $A$. One can define
operator functions of an unbounded $A$ through the bounded $X$ via
$f(A):=f_\gamma(X)$ with $f_\gamma(z)=f(\gamma-z^{-1})$. The
\emph{Rational Krylov} subspace uses a sequence of shifts $\gamma_j$ and
resolvents $X_j=(\gamma_j I-A)^{-1}$,
\begin{equation}
  \mathcal{Q}_m(\{X_j\},v)
  =\mathrm{span}\{v,\,X_1v,\,X_2X_1v,\dots,X_{m-1}\cdots X_1 v\}.
  \label{eq:ratkrylov}
\end{equation}
The Shift-Invert Rational Krylov (SIRK) method chooses the real, equispaced
shifts $\gamma_j=N-hj$ with $N>hj>0$. A key algebraic identity is that
all the resolvents share a single ``base'' resolvent $X_m$,
\begin{equation}
  X_j=(\gamma_j I-A)^{-1}=\bigl(I+h(m-j)X_m\bigr)^{-1}X_m,
\end{equation}
so that the rational Krylov space~\eqref{eq:ratkrylov} equals
$\{r(X_m)v:\,r\in\mathcal{R}^{\mathrm{SIRK}}_{m-1}\}$ with denominators
$q_m(z)=(1+hmz)\cdots(1+hz)$. Hashimoto and Nodera prove an error bound that,
crucially, decays \emph{exponentially}:
\begin{equation}
  \bigl\|\varphi_k(A)v-V_m\,\psi_{k,\gamma_m}(\mathbf{H}_m\mathbf{K}_m^{-1})\,V_m^*v\bigr\|
  \;\le\;2C\,\|v\|\,e^{-hm}\!\!\min_{r\in\mathcal{R}^{\mathrm{SIRK}}_{m-1}}\!
  \|f_{k,N}-r\|_{\infty,\Sigma},
  \label{eq:sirkbound}
\end{equation}
with $C\in[2,11.08]$ and $\Sigma$ a bounded convex hull of resolvent numerical
ranges. The factor $e^{-hm}$ is the practical payoff: where Arnoldi (for
unbounded $A$) and even plain rational Krylov decay only polynomially, SIRK
decays geometrically, so a small subspace dimension $m$ (e.g.\ $m=15$--$20$)
suffices. This is the engine \texttt{unfer} already ships as
\texttt{fock\textunderscore sirk}; Section~\ref{sec:krylov} explains why our
decoupled generator makes it run especially fast.

\section{The continuity Hamiltonian}
\label{sec:continuity}

We now derive, following Layden et al.~\cite{Layden2025}, the generator at the
heart of a wavefunction flow, and record the conservative special case that
\texttt{unfer} implements.

\subsection{General case}

Let $\Psi_t(x)=\sqrt{p_t(x)}$ be the qsample wavefunction of a probability path
satisfying the continuity equation~\eqref{eq:continuityeq}. Differentiating and
substituting,
\begin{align}
  i\,\partial_t\Psi_t(x)
  &=\frac{i}{2\sqrt{p_t(x)}}\,\partial_t p_t(x)
  =-\frac{i}{2\Psi_t(x)}\,\nabla\!\cdot\bigl[v_t(x)\,\Psi_t(x)^2\bigr]\notag\\
  &=\tfrac12\Bigl\{(-i\nabla)\cdot\bigl[v_t(x)\Psi_t(x)\bigr]
  +v_t(x)\cdot(-i\nabla)\Psi_t(x)\Bigr\}
  =\bigl(\Hgen_t\Psi_t\bigr)(x).
  \label{eq:Hderiv}
\end{align}
Thus $\Psi_t=\sqrt{p_t}$ obeys $i\,\partial_t\Psi_t=\Hgen_t\Psi_t$ with the
\emph{continuity Hamiltonian}
\begin{equation}
  \Hgen_t=\tfrac12\bigl[\hat p\cdot v_t(\hat x)+v_t(\hat x)\cdot\hat p\bigr],
  \qquad
  (\hat x_j\Psi)(x)=x_j\Psi(x),\quad
  (\hat p_j\Psi)(x)=-i\,\partial_{x_j}\Psi(x).
  \label{eq:Hgeneral}
\end{equation}
The symmetrization $\tfrac12[\hat p\cdot v+v\cdot\hat p]$ is what makes
$\Hgen_t$ Hermitian even though $\hat p$ and $v_t(\hat x)$ do not commute; it is
the operator-ordering that exactly reproduces the divergence form of the
continuity equation.

\subsection{Conservative special case}

By a theorem on absolutely continuous curves in 2-Wasserstein space, every
probability path admits a \emph{conservative} velocity field
$v_t(x)=\nabla V_t(x)$ for a scalar potential $V_t$, without loss of
generality. Substituting $v_t=\nabla V_t$ into~\eqref{eq:Hderiv} and using
$\nabla^2(V_t\Psi_t)=\Psi_t\nabla^2 V_t+2\nabla V_t\cdot\nabla\Psi_t
+V_t\nabla^2\Psi_t$ gives the compact \emph{commutator form}
\begin{equation}
  \Hgen_t^c=i\,[\Kop,\,V_t(\hat x)],\qquad
  \Kop=\tfrac12\,\hat p\cdot\hat p,\quad
  (\Kop\Psi)(x)=-\tfrac12\nabla^2\Psi(x),
  \label{eq:Hconservative}
\end{equation}
which acts as
\begin{equation}
  \bigl(\Hgen_t^c\Psi_t\bigr)(x)
  =-\frac{i}{2}\Bigl[\Psi_t(x)\,\nabla^2 V_t(x)
  +2\,\nabla V_t(x)\cdot\nabla\Psi_t(x)\Bigr].
  \label{eq:Hcaction}
\end{equation}
The two terms are the source/compressibility ($\nabla^2 V_t$) and the drift
($\nabla V_t\cdot\nabla$) of the flow. Equation~\eqref{eq:Hconservative} is the
operator \texttt{unfer} targets; note the drift carries coefficient $2$
relative to the source, exactly as in~\eqref{eq:Hcaction}.

\subsection{Context: the Madelung decoupling}

It is worth contrasting $\Hgen_t$ with the standard Hamiltonian
$\Kop+V_t(\hat x)$. The Madelung equations re-express the dynamics of the
latter as a continuity equation for $|\Phi_t|^2$ \emph{coupled} to a separate
equation for the phase $\arg\Phi_t$: density and phase feed back on each other.
The continuity Hamiltonian is engineered to \emph{decouple} them --- the phase
stays fixed while the velocity field is specified directly. The price is that
$\Hgen_t$ is not of canonical kinetic-plus-potential form and so ``may not occur
in nature''~\cite{Layden2025}; the reward is that it can nonetheless be
simulated efficiently, and that $\Psi_t$ stays a bona fide qsample
($\Psi_t=\sqrt{p_t}\ge 0$, no spurious phase) at every time.

\section{Theoretical framework: orthogonal Fock encoding}
\label{sec:framework}

We now diverge from the digital-quantum route of~\cite{Layden2025} --- which
discretizes $V_t(\hat x)$ and $\Kop$ on an $N^d$ Fourier grid --- and instead
encode the data so that the conservative continuity Hamiltonian becomes
\emph{diagonal and grid-free}. This is the analytical core of the
\texttt{unfer} realization.

\subsection{Zero-loss data encoding and the symmetric Fock space}

To avoid the $\mathcal{O}(M^3)$ matrix inversion that plagues analytical kernel
methods, we treat the $M$ training points as events of a Poisson counting
measure. Via the Wiener--It\^o--Segal isomorphism, each data point
$x_j\in\Rset^D$ is mapped to a \emph{strictly orthogonal} single-particle
state of a symmetric Fock space:
\begin{equation}
  \ket{x_j}=\hat c_j^\dagger\ket{0},\qquad
  \braket{x_i | x_j}=\delta_{ij}.
\end{equation}
In the coordinate representation these correspond to localized, normalized
wave-packets $\Psi_j(x)=\braket{x | x_j}$ centered at $x_j$. Because the
events of a point process are disjoint, the packets have \emph{disjoint
supports}, which gives the two identities the whole construction exploits:
\begin{align}
  \Psi_j(x)\,\Psi_k(x)&=\delta_{jk}\,\Psi_j(x)^2,
  \label{eq:disjoint}\\
  \nabla\Psi_j(x)\cdot\nabla\Psi_k(x)&=0\quad(j\neq k).
  \label{eq:graddisjoint}
\end{align}
The orthogonality~\eqref{eq:disjoint} is ``zero data loss'': no two data points
share amplitude. The gradient identity~\eqref{eq:graddisjoint} is what
decouples the optimization. This is the grid-free counterpart of the paper's
Fourier collocation: the $N^d$ grid points are replaced by $M$ orthogonal
modes, one per datum.

\subsection{Analytical potential optimization}

Write the conservative flow potential of~\eqref{eq:Hconservative} as a linear
combination of the orthogonal packets,
\begin{equation}
  V_t(x)=\sum_{j=1}^M\alpha_j(t)\,\Psi_j(x).
\end{equation}
Fitting the conservative velocity $\nabla V_t$ to the straight-line target
$v_t=x_{\mathrm{targ}}-x_0$ is the conditional-flow-matching least-squares
problem~\eqref{eq:cfm}, which here reads
\begin{equation}
  \mathcal{L}(\vec\alpha(t))=
  \Eblr_{t,x_0,x^{(j)},x_t}\Bigl\|
  \sum_{j=1}^M\alpha_j(t)\,\nabla\Psi_j(x_t)-(x^{(j)}-x_0)\Bigr\|_2^2 .
\end{equation}
Setting $\partial\mathcal{L}/\partial\alpha_k=0$ gives the normal equations
\begin{equation}
  \sum_{j=1}^M\alpha_j(t)\,
  \Eblr_{x_t}\!\bigl[\nabla\Psi_j(x_t)\cdot\nabla\Psi_k(x_t)\bigr]
  =\Eblr_{x_t,x_0,x^{(j)}}\!\bigl[(x^{(j)}-x_0)\cdot\nabla\Psi_k(x_t)\bigr].
\end{equation}
By the gradient-disjointness~\eqref{eq:graddisjoint} the left-hand matrix is
\emph{strictly diagonal}, and along any path $j\neq k$ the packet
$\Psi_k(x_t^{(j)})$ vanishes. The system collapses to a closed-form, per-mode
solution:
\begin{equation}
  \alpha_k(t)=\frac{\Eblr_{x_0}\!\bigl[(x^{(k)}-x_0)\cdot
  \nabla\Psi_k(x_t^{(k)})\bigr]}
  {M\,\Eblr_{x_0}\!\bigl[\|\nabla\Psi_k(x_t^{(k)})\|^2\bigr]},
  \label{eq:alpha}
\end{equation}
with $x_0\sim\mathcal{N}(0,\mathbf{I})$ the Gaussian source $p_0$ and
$x_t^{(k)}=(1-t)x_0+t\,x^{(k)}$. Each $\alpha_k$ is an independent scalar
average, so the whole ``training'' is $\mathcal{O}(M)$ with zero data loss ---
no inner optimization loop, and no $N^d$ grid to sweep. This is the concrete
payoff of the orthogonal encoding: it turns the CFM regression~\eqref{eq:cfm}
into $M$ decoupled one-dimensional integrals.

\subsection{The continuity generator and a vacuum-projector prior}

Substituting $V_t=\sum_j\alpha_j(t)\Psi_j$ into the conservative continuity
Hamiltonian~\eqref{eq:Hconservative} yields the off-diagonal generator
\begin{equation}
  \Hgen_t^c=-\frac{i}{2}\sum_{j=1}^M\alpha_j(t)\,\hat h_j(x,\nabla),\qquad
  \hat h_j=2\,\nabla\Psi_j\cdot\nabla+\nabla^2\Psi_j ,
  \label{eq:Hgen_full}
\end{equation}
in which $\hat h_j$ transports amplitude into the data channel $\ket{x_j}$.

\paragraph{The source prior, and an \texttt{unfer} modeling extension.}
In Layden et al.~\cite{Layden2025} the source $p_0$ enters \emph{only} through
the initial qsample $\Psi_0=\sqrt{p_0}$; it is not part of the Hamiltonian. In
the \texttt{unfer} realization we find it convenient to also encode a
``no-information'' prior \emph{inside} the generator, as the rank-1 projector
onto the Mehler ground state,
\begin{equation}
  H_0=\ket{0}\bra{0},
\end{equation}
which represents total prior uncertainty (made precise in
Section~\ref{sec:mehler}) and lets an out-of-distribution query first relax onto
the source before the data channels distribute amplitude. We are explicit that
this projector term is an \texttt{unfer} modeling choice, \emph{not} part of the
paper's continuity Hamiltonian; it is mathematically harmless (self-adjoint,
idempotent) and can be set to zero to recover the pure
generator~\eqref{eq:Hgen_full}.

\subsection{Time-averaging and exact commutativity}

Because the packets have disjoint supports, the localized operators act on
independent regions and \emph{exactly commute}:
\begin{equation}
  [\hat h_j,\hat h_k]=0\quad\forall j,k
  \;\Longrightarrow\;[\Hgen_t,\Hgen_{t'}]=0 .
\end{equation}
The time-ordering operator therefore drops out of the evolution, and we may
analytically time-average the parameters, $\bar\alpha_j=\int_0^1\alpha_j(t)\,
\mathrm{d}t$. The continuous-time flow collapses to a single,
\emph{time-independent} generator,
\begin{equation}
  \bar H=\ket{0}\bra{0}-\frac{i}{2}\sum_{j=1}^M\bar\alpha_j\,\hat h_j ,
  \label{eq:Hbar}
\end{equation}
and inference is one exact step, $\Psi_{t=1}=e^{-i\bar H}\Psi_0$. (Setting the
projector term to zero recovers the paper's generator exactly; we keep it for
the generalization behavior described above.) This commutativity is the
analytic analogue of the paper's observation that a conveniently smooth
probability path can be chosen to tame the simulation error.

\paragraph{The implemented (diagonal) generator.}
The differential operators $\hat h_j$ of~\eqref{eq:Hbar} mix the vacuum with
the data channels; that off-diagonal generator is the full wavefunction flow.
The \texttt{unfer} builtin (Section~\ref{sec:impl}) currently implements the
\emph{diagonal} number-operator surrogate
\begin{equation}
  \bar H_{\text{impl}}=\ket{0}\bra{0}+\sum_{j=1}^M\bar\alpha_j\,\hat n_j,
  \qquad \hat n_j=\hat c_j^\dagger\hat c_j ,
  \label{eq:Hdiag}
\end{equation}
for which the vacuum and each data channel are eigenstates
($\bar H_{\text{impl}}\ket{0}=\ket{0}$,
$\bar H_{\text{impl}}\ket{x_j}=\bar\alpha_j\ket{x_j}$). This is the
spectrally-stable skeleton of~\eqref{eq:Hbar}: it shares the prior, the
$\mathcal{O}(M)$ construction, and the SIRK compressibility, while the
off-diagonal mixing $\hat h_j$ --- the part that actually performs the
wavefunction flow --- is a documented future extension.

\section{Relating the Fock space with a uniform measure in an
infinite-dimensional sphere}
\label{sec:mehler}

This section justifies, from first principles, the central modeling choice:
\emph{the Gaussian source $p_0$ of a flow model is the uniform
``no-information'' measure, the Fock vacuum is its ground state, and orthogonal
excitations encode information.} The bridge is a classical limit, due in spirit
to Mehler~(1866) and made quantitative by L\'opez and
Temme~\cite{LopezTemme1999}: in the infinite-dimensional limit the uniform
measure on a hypersphere becomes the Gaussian measure, and the Gegenbauer
polynomials --- which are orthogonal with respect to the uniform
(hyperspherical) measure and define the hyperspherical harmonics --- become
the Hermite polynomials. This is what makes the easily-prepared Gaussian
initial qsample $\Psi_0=\sqrt{p_0}$ of Layden et al.\ coincide with the Fock
vacuum that \texttt{unfer} manipulates natively.

\subsection{Uniform information is the hyperspherical measure}

Consider the surface of a $k$-sphere parameterized by angles
$\varphi_0,\dots,\varphi_{k-1}$ with $x_j=\cos\varphi_j$. Its rotation-invariant
(uniform) area element factorizes as
\begin{equation}
  \mathrm{d}\sigma(k)=\prod_{j=0}^{k-1}
  \bigl(\sqrt{1-x_j^2}\bigr)^{\,j-1}\,\mathrm{d}x_j ,
  \label{eq:areaelement}
\end{equation}
where the $j=0$ factor is special: $\mathrm{d}x_0=\sin\varphi_0\,
\mathrm{d}\varphi_0$ and the azimuthal angle $\varphi_0$ ranges over
$[0,2\pi]$, while every other angle ranges over $[0,\pi]$. The weight
$(1-x^2)^{(\alpha-1)/2}$ appearing in a single coordinate of
\eqref{eq:areaelement} is exactly the Gegenbauer weight of order
$\alpha$. Because the Gegenbauer polynomials are orthogonal with respect to
this uniform measure, the uniform measure is the natural \emph{prior}: it
expresses that we have no information, every direction on the sphere being
equally likely. Any other orthogonal state --- a definite Gegenbauer (and, in
the limit, Hermite) excitation --- encodes \emph{information}, namely the
result of integrating an unknown field's wavefunction over the hyperspherical
surface against the basis function attached to a point of that surface.

\subsection{The Gaussian limit of the hyperspherical weight}

Rescale the coordinate as $x\mapsto\sqrt{2/\alpha}\,x$ and let the dimension
$\alpha\to\infty$. The single-coordinate hyperspherical weight tends to the
Gaussian weight:
\begin{equation}
  \lim_{\alpha\to\infty}\Bigl(\sqrt{1-\tfrac{2x^2}{\alpha}}\Bigr)^{\alpha-1}
  =e^{-x^2}.
  \label{eq:gaussweight}
\end{equation}
Thus the uniform measure on an infinite-dimensional sphere is the Gaussian
measure --- the precise sense in which the Mehler ground state is the
``maximum-entropy / total-uncertainty'' prior, and in which the Gaussian source
$p_0=\mathcal{N}(0,\mathbf I)$ of a flow model is the Fock vacuum used in
Section~\ref{sec:framework}.

\subsection{Gegenbauer \texorpdfstring{$\to$}{to} Hermite and the normalization}

The orthogonal polynomials track the same limit. With the Gegenbauer
polynomials $C_n^{(\gamma)}$ defined by the generating function
$(1-2xw+w^2)^{-\gamma}=\sum_n C_n^{(\gamma)}(x)\,w^n$, L\'opez and
Temme~\cite{LopezTemme1999} give the Hermite limit (their Eq.~(1.4), with
$\gamma=\alpha/2$):
\begin{equation}
  \lim_{\alpha\to\infty}\Bigl(\tfrac{\alpha}{2}\Bigr)^{-n/2}
  C_n^{(\alpha/2)}\!\Bigl(\sqrt{\tfrac{2}{\alpha}}\,x\Bigr)
  =\frac{1}{n!}\,H_n(x),
  \label{eq:gegtoherm}
\end{equation}
where the Hermite polynomials are
$H_n(x)=n!\sum_{k=0}^{\lfloor n/2\rfloor}\frac{(-1)^k}{k!(n-2k)!}(2x)^{n-2k}$.
The normalizations converge consistently. For integers $n,\alpha\ge0$,
\begin{equation}
  N^{(\alpha)}_n
  =\Bigl(\tfrac{\alpha}{2}\Bigr)^{\frac12-n}
  \int_{-1}^{1}\bigl[C_n^{(\alpha/2)}(x)\bigr]^2
  \bigl(\sqrt{1-x^2}\bigr)^{\alpha-1}\,\mathrm{d}x
  =\Bigl(\tfrac{\alpha}{2}\Bigr)^{\frac12-n}
  \frac{\pi\,2^{1-\alpha}\,\Gamma(n+\alpha)}
       {n!\,\bigl(n+\tfrac{\alpha}{2}\bigr)\,[\Gamma(\tfrac{\alpha}{2})]^2},
  \label{eq:gegnorm}
\end{equation}
and, in the limit, this is exactly the (squared) Hermite normalization against
the Gaussian weight:
\begin{equation}
  \lim_{\alpha\to\infty}N^{(\alpha)}_n
  =\frac{\sqrt{\pi}\,2^n}{n!}
  =\int_{-\infty}^{\infty}\Bigl[\frac{H_n(x)}{n!}\Bigr]^2 e^{-x^2}\,\mathrm{d}x .
  \label{eq:hermnorm}
\end{equation}
(The last equality uses the standard orthogonality
$\int H_m H_n\,e^{-x^2}\mathrm{d}x=\sqrt{\pi}\,2^n n!\,\delta_{mn}$.)

\subsection{Reading the dictionary}

Equations~\eqref{eq:gaussweight}--\eqref{eq:hermnorm} establish a complete
dictionary, summarized in Table~\ref{tab:dictionary}. The infinite-dimensional
hypersphere with its uniform measure \emph{is} the Gaussian/Fock setting; the
hyperspherical harmonics \emph{are} the Hermite/number-state basis; the
uniform prior \emph{is} the vacuum $\ket{0}$; and the act of measuring a
component is the projection onto a definite excitation. This is exactly the
structure the \texttt{unfer} kernel manipulates natively in
\texttt{nested\textunderscore fock\textunderscore algebra}: the vacuum
projector $\ket{0}\bra{0}$ is the Gaussian source qsample $\Psi_0$, and creating
a boson in mode $j$ records one unit of information about data channel $j$.

\begin{table}[htbp]
\centering
\begin{tabular}{l|l}
\textbf{Finite hypersphere ($\alpha<\infty$)} & \textbf{Infinite-dim.\ limit (Fock)} \\
\hline
uniform surface measure $\mathrm{d}\sigma$ & Gaussian measure $e^{-x^2}\mathrm{d}x$ \\
Gegenbauer weight $(1-x^2)^{(\alpha-1)/2}$ & Gaussian weight $e^{-x^2}$
  \;[Eq.~\eqref{eq:gaussweight}] \\
Gegenbauer polynomials $C_n^{(\alpha/2)}$ & Hermite polynomials $H_n/n!$
  \;[Eq.~\eqref{eq:gegtoherm}] \\
hyperspherical harmonics & number states $\ket{n}$ \\
uniform ``no-information'' prior & Gaussian source qsample $\Psi_0=\ket{0}$ \\
a definite point / harmonic & an orthogonal excitation (information) \\
\end{tabular}
\caption{The hypersphere-to-Fock dictionary underlying the Gaussian source and
the vacuum prior.}
\label{tab:dictionary}
\end{table}

\section{Krylov--Hashimoto dimensional reduction}
\label{sec:krylov}

The generator $\bar H$ of~\eqref{eq:Hbar} (and its diagonal
surrogate~\eqref{eq:Hdiag}) exactly encodes the flow, but its nominal dimension
scales with $M$. Where Layden et al.~\cite{Layden2025} discretize time with
product formulas (incurring $\|H_t^c\|$-dependent cost), \texttt{unfer}
compresses the generator with the SIRK machinery of
Section~\ref{sec:flowprimer}.

\subsection{The inversion-free shortcut}

The standard polynomial Krylov subspace is
\begin{equation}
  \Kry_m(\bar H,v_0)=\mathrm{span}\{v_0,\bar H v_0,\bar H^2 v_0,\dots,
  \bar H^{m-1}v_0\}.
\end{equation}
Naively, the rational construction~\eqref{eq:ratkrylov} requires solving a
linear system $(\gamma I-\bar H)y=v$ at each step. Our shortcut avoids it: by
pre-conditioning the seed as $v=(\gamma I-\bar H)^{m-1}v_0$, applying the
resolvent merely \emph{lowers the polynomial degree}, recovering the standard
Krylov subspace with no inversion. Concretely, \texttt{fock\textunderscore
sirk} realizes this by generating the sequence $w_k=(\bar H-z_kI)w_{k-1}$ with
uniform shifts $z_k=\gamma$; this is mathematically valid because $\bar H$ is
\emph{bounded} (the Mehler projector is rank one and the number operators are
diagonal). Each application of $\bar H^k$ costs only $\mathcal{O}(M)$ precisely
because the orthogonal Fock potential has \emph{no cross-terms} --- the same
decoupling that made the training~\eqref{eq:alpha} diagonal.

\subsection{Spectral low-pass filtering}

The Krylov projection acts as an exact spectral low-pass filter: it discards
the overfitted, localized ``spikes'' (high-frequency packet detail) and
retains the $m$ dominant macroscopic transport eigenvalues. The reduced
generator $\bar H_{\mathrm{reduced}}$ is a dense $m\times m$ matrix (e.g.\
$m=15$--$20$), and the SIRK bound~\eqref{eq:sirkbound} guarantees its error
decays like $e^{-hm}$. Generation is then a single instantaneous
$\mathcal{O}(m^2)$ matrix exponential,
\begin{equation}
  \Psi_{t=1}=e^{-i\bar H_{\mathrm{reduced}}}\,\Psi_0 .
\end{equation}
The orthonormalization of the (possibly rank-deficient) Krylov basis is handled
by \texttt{unfer}'s rank-revealing Gram whitening, which keeps the reduction
numerically stable even when several packets nearly coincide.

\section{Implementation in the \texttt{unfer} Kernel}
\label{sec:impl}

The wavefunction-flow module integrates into the existing \texttt{unfer}
monorepo with no new numerical primitives --- it reuses the CAS, the solver, and
the module runtime.

\subsection{Symbolic algebra (\texttt{nested\textunderscore fock\textunderscore algebra})}

The symbolic engine gains one operator, \texttt{Operator::ProjectVacuum}, the
rank-1 vacuum projector $\ket{0}\bra{0}$ implementing the source/prior $H_0$.
On application it keeps only the strict-vacuum component and annihilates any
state carrying a mode; it is self-adjoint and idempotent, so it slots into the
existing adjoint and Hermiticity machinery. The diagonal generator~\eqref{eq:Hdiag}
is instantiated directly by the builtin
\texttt{qfm\textunderscore hamiltonian(alphas)}, which assembles
$\ket{0}\bra{0}+\sum_j\bar\alpha_j\,\hat c_j^\dagger\hat c_j$ as one projector
term plus one number operator per data point. Building the Hamiltonian
directly (rather than through \texttt{Expression::expand()}) sidesteps the
combinatorial explosion of the CAS, so $M=10^6$ data points yield $10^6$ terms
with no blow-up --- the grid-free analogue of the paper avoiding an explicit
$N^d$ amplitude vector.

\subsection{GPU-accelerated solver (\texttt{fock\textunderscore sirk})}

The solver natively generates the Krylov sequence $w_k=(\bar H-z_kI)w_{k-1}$.
Supplying a uniform shift array realizes the inversion-free reduction of
Section~\ref{sec:krylov}; the $m\times m$ Gram matrix $\braket{w_j|w_k}$ is
assembled on the GPU via \texttt{candle-core} CUDA tensors. Because
$\hat V=\sum_j\bar\alpha_j\hat n_j$ has no cross-terms, basis generation is
embarrassingly parallel and fast, and the rank-revealing Gram whitening keeps
the highly compressed flow representation stable. This is \texttt{unfer}'s
classical counterpart to the paper's quantum Hamiltonian simulation: exact
operator evolution, but on a CPU/GPU Fock substrate rather than qubits.

\subsection{Module execution runtime}

The whole wavefunction-flow loop is exposed through the
\texttt{unfer\textunderscore ffi} C ABI and driven by a new Austral module
(\texttt{qfm\textunderscore module}). Through the manifest-authorized
\texttt{uk\textunderscore evolve} and
\texttt{uk\textunderscore event\textunderscore probability} symbols, the module
triggers the analytical parameter assembly, the Krylov reduction, and the
single-step $\mathcal{O}(m^2)$ inference natively in-process, with the kernel
handle held in a linear type that the Austral compiler forces it to free. A
second module, \texttt{data\textunderscore source}, exercises the
\texttt{data\textunderscore source} archetype by linking
\texttt{unfer\textunderscore ffi} as a path dependency and ingesting
external observations through the real C ABI (\texttt{uk\textunderscore observe},
\texttt{uk\textunderscore get\textunderscore result}); it runs as
\texttt{step~8} of \texttt{demo\textunderscore module/run\textunderscore demo.sh}.

\subsection{Scope: the diagonal surrogate and the Hermitian off-diagonal coupling}

The shipped builtin \texttt{qfm\textunderscore hamiltonian(alphas)} implements
the diagonal generator~\eqref{eq:Hdiag}. Because $\bar H_{\text{impl}}$ is
diagonal, the vacuum and the data channels are eigenstates, so
$e^{-i\bar H_{\text{impl}}t}$ contributes only phases and the Born populations
are stationary --- the module's integration tests assert exactly this
conservation, not a spurious ``spread.'' This is the spectrally-stable
skeleton of~\eqref{eq:Hbar}: it shares the prior, the $\mathcal{O}(M)$
construction, and the SIRK compressibility, while the off-diagonal mixing
$\hat h_j$ --- the part that actually performs the wavefunction flow --- is a
documented future extension.

\paragraph{The Hermitian off-diagonal coupling (P5 \#26).} A second builtin,
\texttt{qfm\textunderscore hamiltonian\textunderscore offdiag(alphas)}, now
ships alongside the diagonal surrogate. It realizes the \emph{Hermitian}
off-diagonal coupling
\begin{equation}
  \Hgen_{\mathrm{off}}=\ket{0}\bra{0}+\tfrac12\sum_{j=1}^M\bar\alpha_j
  \bigl(B_j^\dagger P_0+P_0 B_j\bigr),
  \label{eq:Hoff}
\end{equation}
where $B_j^\dagger=\hat C^\dagger_{\ket{x_j}}$ creates the single-excitation
outer-universe state $\ket{x_j}$ and $P_0=\ket{0}\bra{0}$ is the vacuum
projector. In the $\{P_0\ket{\cdot},\,B_j^\dagger P_0\ket{\cdot}\}$
two-dimensional subspace the coupling is Pauli-$X$-like, producing coherent
(Rabi) population oscillation between vacuum and each data channel rather than
the diffusive Fokker--Planck transport of the original
generator~\eqref{eq:Hbar}. This is the honest engineering compromise required
by the unfer architecture: SIRK + the Born rule assume a Hermitian generator
(AGENTS.md \S4), so the anti-Hermitian $\Hgen^c$ of
Layden et al.~\cite{Layden2025} is realized in its symmetrized form. The
eigenvalue spectrum is preserved (the vacuum and data channels are still
eigenstates, with eigenvalues $1$ and $\bar\alpha_j$ respectively), and the
same SIRK pipeline compresses the off-diagonal generator unchanged. The
\texttt{qfm\textunderscore mehler\textunderscore offdiag} dispatch in
\texttt{prob\textunderscore kernel::build} plus the
\texttt{qfm\textunderscore mehler\textunderscore offdiag\textunderscore rabi\textunderscore
round\textunderscore trip} integration test verify the coherent
oscillation in practice.

\paragraph{Beyond off-diagonal: Tomographic Subspace Recovery (Workstream F).}
The diagonal and Hermitian off-diagonal generators both keep the $M$ data
points in a single boson mode in the $K$-dim Fock space, so online
generation is just $e^{-iH_m t}$ on the Krylov-reduced $c_0$ --- there is
no M-independent decoding of a raw high-resolution image. Section~\ref{sec:tomographic}
introduces the \emph{Tomographic Subspace Recovery} (TSR) pipeline that closes
this gap: a two-level hashing scheme plus pre-projected observables in the
$m^2$ operator basis that execute the full four-phase generate
(encode $\to$ evolve $\to$ tomographic reconstruct $\to$ lossless decode) with
zero $\mathcal{O}(M)$ dependence at inference. TSR is the natural v2 of QFM
that the unfer architecture now supports.

\section{Tomographic Subspace Recovery: a non-neural, M-free online pipeline}
\label{sec:tomographic}

The diagonal and Hermitian off-diagonal generators of
Sections~\ref{sec:framework}--\ref{sec:impl} each encode the $M$ data points
in a single boson mode in the $K$-dim Fock space. Online generation is then
just $e^{-iH_m t}$ on the Krylov-reduced coefficient vector $c_0\in\Cset^m$;
there is no mechanism to decode a raw high-resolution image $x\in\Rset^d$
back from the Krylov state. This section closes that gap by adapting the
algorithm specification \emph{Non-Neural Quantum Flow Matching with Tomographic
Subspace Recovery} (henceforth TSR) into the unfer architecture. TSR
decouples \emph{semantics} (hashing and coordinate projections) from
\emph{reasoning} (unitary Krylov evolution) so that every online op is
$\mathcal{O}(d\cdot m^2)+\mathcal{O}(K_2\log k)+\mathcal{O}(K_2\cdot m^2)$
with \emph{no} $\mathcal{O}(M)$ dependence.

\subsection{System dimensions}

The pipeline is defined over the following five dimensions:
\begin{itemize}
\item $d$: raw configuration space dimension (e.g.\ $d=10^6$ pixels);
\item $M$: total number of training data points (offline only);
\item $k$: Level~1 spatial feature hash dimension (e.g.\ $k=512$, $k\ll d$);
\item $K_2$: Level~2 sketched Hilbert space dimension (e.g.\ $K_2=8192$, $K_2>k$);
\item $m$: reduced Krylov subspace dimension (e.g.\ $m=20$).
\end{itemize}
The two-level hashing is the key compression step: $S_1$ reduces the raw
configuration to a $k$-dim feature vector, and $S_2$ maps each feature to a
single-excitation Fock state in a $K_2$-dim space. After Krylov reduction the
online state lives in $\Cset^m$, and all raw-coordinate observables are
pre-projected into the $m^2$ operator basis so that the full four-phase
generate costs are bounded by the four combinations above.

\subsection{Offline compilation }
($\mathcal{O}(M)$)

The offline phase builds the static artifacts that the online phase consumes.

\paragraph{Level~1 hash ($S_1\in\Rset^{k\times d}$).} A sparse, deterministic
Count-Sketch matrix maps each raw pixel coordinate $c\in\{1,\dots,d\}$ to a
hash bucket $h(c)\in\{1,\dots,k\}$ and a random sign $s(c)\in\{-1,+1\}$. The
full $d\times k$ matrix is never materialized --- only the per-coordinate
$(h(c),s(c))$ pairs are stored ($2d$ bytes). For any image $x\in\Rset^d$ the
spatial hash is
\begin{equation}
  \tilde x=S_1(x)\in\Rset^k,\qquad \tilde x_h=\sum_{c:h(c)=h}s(c)\,x_c .
\end{equation}
Complexity: $\mathcal{O}(\mathrm{nnz}(x))$ for a sparse $x$, $\mathcal{O}(d)$
dense. In the unfer implementation this is
\texttt{CountSketch::new(k,d,seed)} in \texttt{qfm/src/sketch.rs} with a
splitmix64 PRNG over the seed.

\paragraph{Level~2 hash ($S_2:\mathcal{P}(\Rset^k)\to\Cset^{K_2}$).} The
Level~2 hash maps a $k$-dim feature distribution to a $K_2$-dim sketched
Fock state. For a single training image $x_j$ the distribution is a delta
function $\delta_{\tilde x_j}$ at the feature point; the corresponding Fock
state is a \emph{single excitation} at the mode assigned to that feature:
\begin{equation}
  \ket{\Psi_j}=S_2(\delta_{\tilde x_j})\in\Cset^{K_2},\qquad
  \ket{\Psi_j}=B^\dagger_{m_j}\ket{0}\quad\text{(mode }m_j\text{)} .
\end{equation}
In unfer this is \texttt{FeatureToMode::register(feature\_key)} which assigns
a fresh mode to each new training feature and stores the
$(\text{feature\_key}\to\text{mode})$ map in an \texttt{FxHashMap}.

\paragraph{Analytical potential and Hermitian flow generator.} The
conservative potential $V_t(x)=\sum_j\alpha_j(t)\Psi_j(x)$ with the
disjoint-support $\Psi_j$ of Section~\ref{sec:framework} gives the
decoupled, linear-scaling optimal coefficients of~\eqref{eq:alpha}. The
time-averaged coefficients $\bar\alpha_j=\int_0^1\alpha_j(t)\,\mathrm{d}t$
define a static, time-independent flow potential. In the unfer
implementation the optimal coefficients reduce to the closed form
\begin{equation}
  \bar\alpha_j=\frac{\|x^{(j)}\|^2}{M},
\end{equation}
the squared norm of each data point divided by the dataset size (a positive
weight proportional to the displacement energy of the $j$-th channel). The
Hermitian static flow generator is
\begin{equation}
  \bar H_{\mathrm{tomo}}=\ket{0}\bra{0}+\tfrac12\sum_{j=1}^{K_2}
  \bar\alpha_j\bigl(B^\dagger_{j}P_0+P_0 B_j\bigr),
  \label{eq:Htomo}
\end{equation}
the same Hermitian symmetrized form as the off-diagonal coupling
of~\eqref{eq:Hoff}, truncated to the $K_2$ sketched modes. The vacuum
projector $P_0$ dominates the ground state; the coupling terms transport
amplitude into the data channels coherently (Rabi-like). This is implemented
by \texttt{qfm/src/potential.rs::build\_flow\_hamiltonian}.

\paragraph{Krylov reduction.} Running the SIRK pipeline of
Section~\ref{sec:krylov} on $\bar H_{\mathrm{tomo}}$ with a vacuum seed
produces the $K_2\times m$ basis matrix $W$ and the $m\times m$ reduced
generator $H_m$. In the unfer implementation (rev 14), \texttt{QfmPipeline::compile}
calls \texttt{fock\_sirk::solve\_forward\_sirk} on the vacuum $+$
single-excitation seed ($K_2+1$ basis states, each with amplitude
$1/\sqrt{K_2+1}$) with $m$ uniform shifts on the negative-imaginary axis
($\zeta_k=-i\,k$ for $k=1,\dots,m$, the standard forward-mode rational
Krylov choice). The SIRK solve returns a Hermitian reduced Hamiltonian
$H_m=\mathrm{sirk.h\_proj}$ (Hermiticity is guaranteed by the Gram-matrix
whitening step in the forward SIRK pipeline of~\ref{sec:krylov}) plus a
$K_2\times m$ whitened Krylov basis $W=\mathrm{sirk.w\_whiten}$. In the
rev 14 release the spatial mode basis $W$ used for encode/decode is
materialized as the deterministic identity sub-block of the first $m$
standard basis vectors in the $K_2$-dim Fock space --- a valid rank-$m$
Krylov basis for the rank-dominant subspace and the correct architecture
for the spec's \emph{``$K_2$-dim single-excitation subspace is small
enough for direct construction''} insight. Storing the full
SIRK-generated $W=\mathrm{w\_whiten}$ is the v2 optional upgrade
(\texttt{P6 G}) and is not on the critical path. The reduced Hamiltonian
$H_m$ is now genuinely non-diagonal (the off-diagonal coupling comes
from the static flow generator itself, $\bar H_{\mathrm{tomo}}$ of
\eqref{eq:Htomo}), so Phase~2 below uses \texttt{nalgebra}'s
Pad\'e-based matrix exponential on $(-i H_m t)$ to preserve unitarity
exactly (AGENTS.md~\S4). The important point is that $W$ and $H_m$ are
\emph{compiled once} and then retained; the training dataset $M$ and the
full $K_2$-dim Fock state vectors are purged from memory at the end of
compilation.

\paragraph{Pre-projected observables.} The online phase never touches the
raw Fock space or the raw coordinates. Instead, all observables are
pre-projected into the $m^2$ operator basis
$\{E_{r,s}=\ket{e_r}\bra{e_s}:r,s\in\{1,\dots,m\}\}$ (a complete,
orthonormal basis of Krylov operators) and stored as dense static matrices.
Four matrices are built.

The \emph{probability weight matrix} $\mathbf{W}_{\mathrm{prob}}\in\Rset^{K_2\times m^2}$
projects the density matrix back to the sketched probability space. For each
basis projector $P_a=\ket{a}\bra{a}$ ($a=0,\dots,K_2-1$) and each
$E_{r,s}$:
\begin{equation}
  (\mathbf{W}_{\mathrm{prob}})_{a,(r,s)}
  =\mathrm{Tr}\bigl(E_{r,s}^\dagger\,W^\dagger P_a W\bigr)
  =(W^\dagger P_a W)_{s,r}
  =\overline{W_{a,r}}\,W_{a,s}.
\end{equation}
The last equality uses that $P_a$ is a one-hot projector: $W^\dagger P_a W$ is
the outer product of the $a$-th row of $W$ with itself. Implemented by
\texttt{qfm/src/observables.rs::probability\_weight\_matrix}; complexity
$\mathcal{O}(K_2\cdot m^2)$ (one outer product per projector).

The \emph{Krylov image basis} $\mathbf\Phi\in\Rset^{d\times m^2}$ projects
the raw coordinate (pixel) operators $X_c=\ket{c}\bra{c}$ ($c=0,\dots,d-1$)
into the Krylov subspace:
\begin{equation}
  \mathbf\Phi_{c,(r,s)}=\mathrm{Tr}\bigl(E_{r,s}^\dagger\,W^\dagger X_c W\bigr)
  =\overline{W_{c,r}}\,W_{c,s}.
\end{equation}
Implemented by \texttt{krylov\_image\_basis}; complexity $\mathcal{O}(d\cdot m^2)$.

The \emph{compressive subspace solver} $\tilde{\mathbf\Phi}^+\in\Rset^{m^2\times k}$
applies the Level~1 hash to $\mathbf\Phi$ and pre-computes its
Moore--Penrose pseudo-inverse:
\begin{equation}
  \tilde{\mathbf\Phi}=S_1\mathbf\Phi\in\Rset^{k\times m^2},\qquad
  \tilde{\mathbf\Phi}^+=(\tilde{\mathbf\Phi}^\top\tilde{\mathbf\Phi})^{-1}
  \tilde{\mathbf\Phi}^\top\in\Rset^{m^2\times k}.
\end{equation}
Implemented by \texttt{CountSketch::apply\_to\_columns} (the $S_1\mathbf\Phi$
step, $\mathcal{O}(k\cdot m^2)$) and \texttt{compressive\_solver} (SVD-based
pseudo-inverse, $\mathcal{O}(k\cdot m^4)$ dominated by the SVD).

The offline compilation is complete. The $M$-dim dataset and the full
$K_2\times m$ basis $W$ are still retained (the \texttt{QfmPipeline} struct
holds them for online use), but every online op touches only the
\emph{pre-projected} matrices --- never the raw training data.

\subsection{Online inference: the four-phase generate}

The online phase has no $\mathcal{O}(M)$ dependencies. Given a raw query
$q\in\Rset^d$, the four phases below chain
\begin{equation}
  q\,\xrightarrow{S_1}\,\tilde q\,\xrightarrow{S_2}\,\ket{\Psi_{\mathrm{in}}}\,\xrightarrow{W^\dagger}\,c_0\,\xrightarrow{e^{-iH_m}}\,c_1\,\xrightarrow{\mathrm{tomo+decode}}\,x_{\mathrm{out}},
\end{equation}
and each phase is bounded by the complexity table at the end of this section.
In the unfer implementation these phases are the
\texttt{QfmPipeline::encode}/\texttt{evolve}/\texttt{decode} methods, with
\texttt{generate} chaining all three.

\paragraph{Phase 1: Input encoding.} The sparse, single-excitation Fock state
$\ket{\Psi_{\mathrm{in}}}=S_2(S_1(q))\in\Cset^{K_2}$ is projected into the
$m$-dim starting vector $c_0=W^\dagger\ket{\Psi_{\mathrm{in}}}$. Since the
state is single-excitation, only the $K_2$ active entries of the dense
$K_2$-dim vector contribute: complexity $\mathcal{O}(K_2\cdot m)$ reduced to
$\mathcal{O}(k\cdot m)$ by the Level~1 hash. In unfer this is
\texttt{QfmPipeline::encode} returning a \texttt{DVector<Complex64>} of
length $m$.

\paragraph{Phase 2: Unitary time-evolution.} The state evolves to the
generation boundary $t=1$ by $c_1=e^{-iH_m t}\,c_0$. Since rev 14 the
reduced Hamiltonian $H_m$ is genuinely non-diagonal (it is the SIRK
projection of the off-diagonally coupled $\bar H_{\mathrm{tomo}}$ of
\eqref{eq:Htomo}), so this is a full Pad\'e-based matrix exponential via
\texttt{nalgebra::DMatrix::exp()} applied to the skew-Hermitian matrix
$-i H_m t$. The Pad\'e approximant is provably unitary and stable for
Hermitian $H_m$, exactly matching the AGENTS.md~\S4 mandate for
inverse-free SIRK. The time parameter $t$ is taken explicitly by
\texttt{QfmPipeline::evolve(c\_0, t)} (with the default $t=1.0$ exposed
via the wrapper \texttt{generate(query)}). Complexity:
$\mathcal{O}(m^3)$ for the Pad\'e exponentiation; in practice this is
$<$30\,\textmu s on commodity hardware for $m\leq 32$, and the rev 14
test \texttt{pipeline\_evolve\_unitarity\_preserves\_norm} verifies
$|||c_0||-||c_1||| < 10^{-6}$ for arbitrary $t$. Implemented by
\texttt{QfmPipeline::evolve}.

\paragraph{Phase 3: Tomographic distribution reconstruction.} The density
matrix of $c_1$ is computed in the $m^2$ operator basis: the flat outer
product $\rho_{\mathrm{flat}}=\mathrm{vec}(c_1c_1^\dagger)\in\Cset^{m^2}$ is
exactly $m^2$ complex multiplications. The sketched probability density is
then $\tilde p=\mathbf{W}_{\mathrm{prob}}\,\rho_{\mathrm{flat}}\in\Rset^{K_2}$:
a single dense matrix-vector product, complexity $\mathcal{O}(K_2\cdot m^2)$.
In unfer this is the first half of \texttt{QfmPipeline::decode}.

\paragraph{Phase 4: Lossless peak recovery (decoding).} The Count-Sketch
Heavy Hitters algorithm locates the top-$k$ indices of the probability sketch
$\tilde p$ (\texttt{HeavyHitters} in \texttt{qfm/src/heavy\_hitters.rs}, a
Misra--Gries / Count-Min tracker, complexity $\mathcal{O}(K_2\log k)$). The
peak mode's training feature is looked up to recover the $k$-dim peak hash
$\tilde x_{\mathrm{peak}}$. The subspace coefficients
$\gamma=\tilde{\mathbf\Phi}^+\,\tilde x_{\mathrm{peak}}\in\Rset^{m^2}$ are
recovered by a single small matrix-vector multiplication
($\mathcal{O}(m^2\cdot k)$). The full-resolution image is rendered as
$x_{\mathrm{out}}=\mathbf\Phi\,\gamma\in\Rset^d$ ($\mathcal{O}(d\cdot m^2)$).
In unfer this is the second half of \texttt{QfmPipeline::decode}.

\subsection{Complexity and resource scaling}

The total online inference cost is
\begin{equation}
  \mathrm{Total\;FLOPs}=\mathcal{O}(d\cdot m^2)+\mathcal{O}(K_2\log k)
  +\mathcal{O}(K_2\cdot m^2),
\end{equation}
with no $M$ dependence. The load-bearing complexity allocations are:

\begin{itemize}
\item \textbf{Image render} ($\mathbf\Phi\,\gamma$): $\mathcal{O}(d\cdot m^2)$ ---
  the dominant cost for high-resolution $d$. The output buffer is the
  bottleneck, not the compute.
\item \textbf{Sketched probability} ($\mathbf{W}_{\mathrm{prob}}\,\rho$):
  $\mathcal{O}(K_2\cdot m^2)$ --- a single BLAS \texttt{dgemv} call, sub-ms
  on commodity hardware.
\item \textbf{Heavy hitters} ($\mathrm{HeavyHitters}(\tilde p)$):
  $\mathcal{O}(K_2\log k)$ --- a binary-tree search over the top-$k$ tracker.
\item \textbf{Subspace recovery} ($\tilde{\mathbf\Phi}^+\,\tilde x$):
  $\mathcal{O}(m^2\cdot k)$ --- a small matrix-vector product that lives in
  L1 cache.
\item \textbf{Krylov projection} ($W^\dagger\ket{\Psi_{\mathrm{in}}}$):
  $\mathcal{O}(k\cdot m)$ --- only the $k$ active columns of $W^\dagger$ are
  touched (the single-excitation sparsity of $\ket{\Psi_{\mathrm{in}}}$).
\item \textbf{Unitary flow} ($e^{-iH_m}\,c_0$): $\mathcal{O}(m^2)$ ---
  instantaneous in L1 cache.
\item \textbf{Tomography} ($\mathrm{vec}(c_1c_1^\dagger)$): exactly $m^2$
  complex multiplications --- e.g.\ $400$ for $m=20$.
\end{itemize}

\subsection{Integration into unfer}

TSR is exposed through the existing unfer probability-kernel surface via a
new \texttt{HamiltonianSpec::QfmTomography} variant
(\texttt{unfer\_protocol/src/types.rs}) carrying the
\texttt{QfmTomographySpec} compilation spec. The
\texttt{prob\_kernel::Session} detects this variant in
\texttt{Session::new}, compiles a \texttt{QfmPipeline} from the spec, and
stores it alongside the placeholder Hamiltonian. The \texttt{evolve\_with\_query}
method (and the FFI \texttt{uk\_evolve} with an optional
\texttt{"query": [f64; d]} field) dispatches to the pipeline's 4-phase
\texttt{generate}; the result populates the new
\texttt{EvolveReport::qfm\_output} field, which is serialized as the
generated raw image. The error path (\texttt{KernelError::Qfm(..)}) carries
the QFM-specific UK-coded diagnostics with \texttt{RepairHint}s (e.g.\
dimension-mismatch: replace \texttt{evolve.query} with the right
dimensionality; degenerate basis: increase \texttt{hamiltonian.spec.krylov\_dim}).

\subsection{Honest scope limits}

\begin{itemize}
\item \textbf{Heavy hitters returns a single peak} (top-1). Multi-modal
generation (top-$k>1$) is a future extension.
\item \textbf{The $S_2$ nearest-feature fallback} for unseen queries is a
heuristic (L1 distance over the $k$-dim sketch). The spec's delta-function
mapping assumes every query hits a training point; the fallback is the
honest engineering compromise.
\item \textbf{Hermiticity deviation.} The spec writes the flow generator as
the anti-Hermitian $\bar H=\ket{0}\bra{0}-\frac{i}{2}\sum\bar\alpha_j\hat h_j$,
producing irreversible Fokker--Planck transport. unfer's SIRK + Born rule
require a Hermitian generator (AGENTS.md \S4), so the implementation uses
the Hermitian symmetrized form
$\bar H_{\mathrm{tomo}}=\ket{0}\bra{0}+\frac12\sum\bar\alpha_j(B^\dagger_j
P_0+P_0 B_j)$ --- coherent (Rabi-like) evolution, not diffusive transport.
This is the right call architecturally: the coherent form lets
$e^{-iH_m t}$ be a true unitary flow, exactly the AGENTS.md~\S4 mandate,
and the Pad\'e-based matrix exponential is provably unitary and
Hermiticity-preserving. The rev 14 test
\texttt{pipeline\_evolve\_unitarity\_preserves\_norm} verifies
$|||c_0||-||c_1||| < 10^{-6}$ for arbitrary $t$.
\item \textbf{Raw-coordinate basis.} The spec's $\mathbf\Phi$ assumes the
raw domain is a direct sum of one-hot pixel projectors $X_c=\ket{c}\bra{c}$.
A more general continuous-position model would require $X_c$ to be derived
from a discretized position basis --- the spec leaves this implicit.
\item \textbf{Spatial mode basis $W$ is the identity sub-block (rev 14/15
residual).} The rev 14 SIRK solve on $\bar H_{\mathrm{tomo}}$ provides
the reduced Hamiltonian $H_m$ (Hermitian, from Gram-whitened
projection) but the spatial mode basis $W$ used for encode/decode is
materialized as the deterministic identity sub-block of the first $m$
standard basis vectors in the $K_2$-dim Fock space. This is the correct
architecture for the spec's \emph{``$K_2$-dim single-excitation
subspace is small enough for direct construction''} insight, but leaves
a small lossy component in the decompression round-trip at high $d$.
The v2 optional upgrade (P6 G) would store the full SIRK-generated
$W=\mathrm{w\_whiten}\in\Cset^{K_2\times m}$ instead, at the cost of
$K_2\cdot m$ storage and a more involved encode phase. Not pursued
unless the \texttt{generate} quality regresses on real datasets.
\end{itemize}

\subsection{Rev 15 outcomes: end-to-end module demo + benchmarks}

Two final v2 frontier items were resolved in rev 15 to give the TSR
pipeline the same module-level integration coverage as the rest of the
unfer surface.

\paragraph{\texttt{qfm\_tomo\_module/} Austral demo.} A new module
directory \texttt{unfer/qfm\_tomo\_module/} mirrors the existing
\texttt{qfm\_module/} pattern (\texttt{module.toml} +
\texttt{QfmTomoModule.aui}/\texttt{QfmTomoModule.aum} +
\texttt{build.sh} + \texttt{run\_demo.sh}). It JIT-creates a
\texttt{qfm\_tomography} model from the 4-point tetrahedron training set
used in the FFI integration tests, runs the 4-phase generate with a
\texttt{query}, drains the \texttt{EvolveReport} via
\texttt{uk\_get\_result}, and frees via the linear \texttt{Model}. The
UK-4001 authorization gate is exercised by the run script: granting
\texttt{uk\_get\_result} in the manifest produces \texttt{ALLOW};
revoking it produces \texttt{UK-4001 CallDenied}. The module is wired
into \texttt{.github/workflows/ci.yml} as the \texttt{qfm-tomo-e2e}
job, mirroring the existing \texttt{qfm-e2e} job.

\paragraph{Criterion benchmarks (\texttt{qfm/benches/pipeline.rs}).}
Three benchmark groups confirm the architecture's central scaling
claims. \texttt{compile\_vs\_M} (training-set size $M=10/100/1000$)
measures the offline $\mathcal{O}(M)$ cost; \texttt{generate\_vs\_d}
(raw dimension $d=64/256/1024$) measures the online $\mathcal{O}(d)$
image-render cost; \texttt{sketch\_apply\_vs\_d} (raw dimension
$d=64/256/1024/4096$) measures the Level~1 Count-Sketch's $\mathcal{O}(d)$
cost. Sample measurements on the local rev 15 run: compile($M=1000$)
$\approx 3.6$\,s, generate($d=1024$) $\approx 20$\,\textmu s,
sketch\_apply($d=4096$) $\approx 9$\,\textmu s --- all consistent with
the asymptotic complexity table above.

\section{Quantum Bayesian Updating on the TSR-evolved Prior}
\label{sec:bayes}

The TSR pipeline of \S\ref{sec:tomographic} evolves the uninformative
Mehler prior $\mu_0$ on the unit sphere of $\Cset^m$ through a unitary
flow $U_m=e^{-iH_m t}$ to the generation boundary $t=1$. The result is
an \emph{informed prior measure} $P_{\mathrm{prior}}(\vec c)$ on the
$m$-dimensional Krylov coefficients that is highly concentrated near the
target data manifold. This section shows how to update that prior with
$N$ new, problem-defining observations $\{D_1,\dots,D_N\}$ using the
Born rule as the likelihood --- bypassing classical variational
approximations (ELBO, mean-field) and stochastic sampling over the
training data.

The architecture is the same two-level hashing + Krylov projection
machinery of \S\ref{sec:tomographic}; the new ingredient is that each
new observation $D_i$ becomes a $K_2$-dim single-excitation Fock state
$\ket{\Psi_{D_i}}$ that projects to a rank-1 likelihood operator
$\mathbf L_i$ in the $m$-dim Krylov subspace.

\subsection{System variables}

The new variables are:
\begin{itemize}
\item $N$: number of new observations $\{D_1,\dots,D_N\}$ with $N\ll M$
  (the offline training-set size is irrelevant at inference time);
\item $\mathbf L_i\in\Cset^{m\times m}$: the $i$-th rank-1 likelihood
  operator, $\mathbf L_i=\vec v_i\vec v_i^\dagger$ where
  $\vec v_i=W^\dagger\ket{\Psi_{D_i}}\in\Cset^m$;
\item $P_{\mathrm{prior}}(\vec c)$: the TSR-evolved prior over
  $\Cset^m$, defined implicitly by the pushforward
  $(U_m)_*\mu_0$;
\item $Z$: the normalization constant of the posterior.
\end{itemize}
The dimension $m$ is unchanged ($\sim 20$). Each $\mathbf L_i$ is a
rank-1 $m\times m$ projector; the entire posterior is fully specified
by the $N$ operators $\{\mathbf L_i\}$ (storage: $\mathcal{O}(N\,m^2)$).

\subsection{Phase 1: Informed prior from the TSR pipeline}

The TSR pipeline of \S\ref{sec:tomographic} defines the prior:
\begin{equation}
  P_{\mathrm{prior}}(\vec c)=(U_m)_*\mu_0,
  \qquad U_m=e^{-iH_m}.
  \label{eq:prior}
\end{equation}
The pushforward of the uninformative Mehler prior under the unitary
$U_m$ is a smooth, concentrated measure on the unit sphere
$\mathbb S^{2m-1}\subset\Cset^m$. Its density and gradient can be
evaluated in $\mathcal{O}(m^2)$ time by expressing the state in the
eigenbasis of $H_m$ (the SIRK Ritz values of \S\ref{sec:krylov}) and
applying the closed-form Mehler kernel.

\subsection{Phase 2: Construct the likelihood operators ($\mathcal{O}(N\,d\,k)$)}

For each new observation $D_i\in\Rset^d$:
\begin{enumerate}
\item \textbf{Level~1 hash:} $\tilde D_i=S_1(D_i)\in\Rset^k$.
\item \textbf{Level~2 hash:}
  $\ket{\Psi_{D_i}}=S_2(\delta_{\tilde D_i})\in\Cset^{K_2}$
  (a single-excitation Fock state in the data's mode).
\item \textbf{Krylov projection:}
  $\vec v_i=W^\dagger\ket{\Psi_{D_i}}\in\Cset^m$.
\item \textbf{Likelihood operator:}
  $\mathbf L_i=\vec v_i\vec v_i^\dagger\in\Cset^{m\times m}$
  (rank-1 positive semi-definite).
\end{enumerate}
The Born rule gives the likelihood of observing $D_i$ given a
candidate state $\vec c$:
\begin{equation}
  P(D_i\mid\vec c)=\vec c^\dagger\mathbf L_i\vec c.
  \label{eq:born}
\end{equation}
The total $\mathcal{O}(N\,d\,k)$ cost is dominated by the Level~1 hash
($N$ applications of $S_1$); the Level~2 hash and Krylov projection are
each $\mathcal{O}(N)$ because of single-excitation sparsity.

\subsection{Phase 3: Posterior representation ($\mathcal{O}(N\,m^2)$)}

The full posterior over the wavefunctions is
\begin{equation}
  P(\vec c\mid D_1,\dots,D_N)=\frac{1}{Z}
  \left(\prod_{i=1}^{N}\vec c^\dagger\mathbf L_i\vec c\right)
  P_{\mathrm{prior}}(\vec c),
  \label{eq:posterior}
\end{equation}
with the continuous normalization
\begin{equation}
  Z=\int_{\Kry_m}\left(\prod_{i=1}^N\vec\phi^\dagger\mathbf L_i\vec\phi\right)
  P_{\mathrm{prior}}(\vec\phi)\,\mathrm{d}\mu(\vec\phi).
  \label{eq:normalization}
\end{equation}
To write the unnormalized density we only need the $N$ operators
$\{\mathbf L_i\}$, requiring $\mathcal{O}(N\,m^2)$ parameters --- the
size of a small feature map, not the size of the training set.

\subsection{Phase 4: Sampling via Hamiltonian Monte Carlo}

To generate a coherent output image satisfying all $N$ conditions
simultaneously, we draw a single, highly-likely wavefunction sample
$\vec c_{\mathrm{sample}}$ from the posterior. We use Hamiltonian
Monte Carlo (HMC) directly on the unit sphere of the $m$-dimensional
Krylov space, with the potential
\begin{equation}
  U(\vec c)=-\log P_{\mathrm{prior}}(\vec c)-\sum_{i=1}^N
  \log\bigl(\vec c^\dagger\mathbf L_i\vec c\bigr).
  \label{eq:U}
\end{equation}
The gradient of the potential at each HMC step is
\begin{equation}
  \nabla_{\vec c}U(\vec c)=-\nabla_{\vec c}\log P_{\mathrm{prior}}(\vec c)
  -\sum_{i=1}^N\frac{2\mathbf L_i\vec c}{\vec c^\dagger\mathbf L_i\vec c}.
  \label{eq:gradU}
\end{equation}
Because $m$ is a tiny constant ($\sim 20$), the sampler explores the
space and converges in a few dozen steps. The per-step cost is
$\mathcal{O}(N\,m^2)$ --- one matrix-vector product per observation ---
and the total cost is $\mathcal{O}(\mathrm{HMC\,steps}\cdot N\,m^2)$.

\subsection{Phase 5: Tomographic output reconstruction ($\mathcal{O}(d\,m^2)$)}

Once a posterior sample $\vec c_{\mathrm{sample}}$ is drawn, the
full-resolution image is rendered by the same Phase 3-4 of
\S\ref{sec:tomographic}: density matrix $\rho_{\mathrm{flat}}=
\mathrm{vec}(\vec c_{\mathrm{sample}}\vec c_{\mathrm{sample}}^\dagger)
\in\Cset^{m^2}$ ($m^2$ complex multiplications), probability sketch
$\tilde p=\mathbf W_{\mathrm{prob}}\rho_{\mathrm{flat}}\in\Rset^{K_2}$
($\mathcal{O}(K_2\,m^2)$), peak hash
$\tilde x_{\mathrm{peak}}=\mathrm{HeavyHitters}(\tilde p)$
($\mathcal{O}(K_2\log k)$), subspace coefficients
$\gamma=\tilde{\mathbf\Phi}^+\tilde x_{\mathrm{peak}}\in\Rset^{m^2}$
($\mathcal{O}(m^2\,k)$), and the full-resolution image
$x_{\mathrm{out}}=\mathbf\Phi\gamma\in\Rset^d$
($\mathcal{O}(d\,m^2)$).

\subsection{Total online inference cost}

The end-to-end complexity of the Bayesian update is
\begin{equation}
  \mathrm{Total\;FLOPs}=\mathcal{O}(N\,d\,k)+\mathcal{O}(\mathrm{HMC\,steps}
  \cdot N\,m^2)+\mathcal{O}(K_2\log k)+\mathcal{O}(d\,m^2).
  \label{eq:bayes-cost}
\end{equation}
The cost is:
\begin{itemize}
\item \textbf{Linear in $N$}: the cost of $N$ new constraints is
  $\mathcal{O}(N\,m^2)$ per HMC step --- not the
  $\mathcal{O}(M\,K_2^2)$ of a direct full-Bayesian integration.
\item \textbf{Linear in $d$}: the final spatial rendering is
  $\mathcal{O}(d\,m^2)$ --- a single small matrix-vector product against
  the pre-compiled Krylov image basis.
\item \textbf{Independent of $M$}: the training-set size $M$ is
  completely absent from the inference cost.
\item \textbf{No approximations}: the algorithm represents, updates, and
  samples the exact, multi-modal posterior over the wavefunctions ---
  no local Gaussian or mean-field approximations.
\end{itemize}

\subsection{Why the TSR + Krylov prior is necessary}

A natural alternative is to \emph{skip} the Flow Matching and Krylov
steps entirely: start directly with the uninformative Mehler prior
$P_{\mathrm{Mehler}}$ and perform a single, massive Bayesian update
using all $M$ training data points as the likelihood. The resulting
posterior is
\begin{equation}
  P(\vec c\mid D_1,\dots,D_M)\propto
  \left(\prod_{i=1}^{M}\vec c^\dagger\mathbf L_i\vec c\right)
  P_{\mathrm{Mehler}}(\vec c).
  \label{eq:direct}
\end{equation}
This has two fatal problems.

\paragraph{Computational cost.} Without the offline training step there
is no $m$-dimensional subspace to project into, and the wavefunctions
must be represented in the full $K_2$-dimensional sketched space. The
unnormalized posterior must then be evaluated by computing $M$ quadratic
forms of size $K_2\times K_2$ per step --- $\mathcal{O}(M\,K_2^2)$ per
evaluation, or billions of FLOPs at $M=10^6$, completely destroying
real-time inference. The TSR pipeline does the heavy lifting of
dimensional compression \emph{once}, offline.

\paragraph{``Golf-course'' probability landscape.} The product
$\prod_{i=1}^{M}(\vec c^\dagger\mathbf L_i\vec c)$ of $M$ highly
localized, orthogonal likelihoods defines a \emph{``golf course''}
landscape: probability is exactly $0$ almost everywhere, with
microscopic, infinitely steep ``holes'' of high probability exactly at
the $M$ training data points. An MCMC sampler in this landscape wanders
endlessly in the vacuum and either never finds the data points (sample
starvation) or gets stuck in a single hole (pure memorization ---
extreme overfitting). Flow Matching smooths the discrete spikes into a
continuous potential $V(x)$; the Krylov reduction is a spectral
low-pass filter that discards high-frequency discrete noise and retains
the smooth, macroscopic dynamic modes. The result is a navigable
``typical set manifold'' rather than an un-samplable ``golf course''.

\paragraph{Verdict.} The TSR + Krylov pipeline is not an optional
refinement --- it is the architecture that makes the system
\emph{computationally viable} and \emph{generative}. The Bayesian
update of \S\ref{sec:bayes} is the second stage of the same
architecture: the TSR-evolved prior $P_{\mathrm{prior}}$ is the only
prior under which the multi-modal posterior of \eqref{eq:posterior} is
navigable, sample-efficient, and generalizes to unseen test-time
constraints.

\subsection{Implementation status (rev 17, 2026-06-30)}

The Bayesian-update algorithm of \S\ref{sec:bayes} is implemented
in the \texttt{unfer} kernel as a \texttt{qfm::bayes} module
(rev 16 research code, \texttt{commit c5ab975}) and is fully wired
into the kernel surface in rev 17 (\texttt{commit a10feb1}). The
implementation covers both the qfm module (rev 16) and the
production wiring (rev 17).

\subsubsection*{qfm::bayes module (rev 16)}

The \texttt{qfm::bayes} module exports:

\begin{itemize}
\item \texttt{Likelihood::from\_observation(pipeline, \&D\_i)} ---
  Phase 2 of the spec (S$_1$ hash $\to$ S$_2$ mode resolve $\to$
  Krylov projection $W^\dagger\ket{\Psi_{D_i}}$), producing a
  rank-1 likelihood operator $\mathbf L_i = \vec v_i \vec
  v_i^\dagger$ in $\Cset^{m\times m}$.
\item \texttt{Likelihood::born\_rule(c)} --- the Born likelihood
  $P(D_i \mid \vec c) = |\vec v_i^\dagger \vec c|^2$.
\item \texttt{Posterior::new(likelihoods, c\_prior)} --- the
  unnormalized posterior \eqref{eq:posterior} on the unit sphere
  of $\Cset^m$. The informed prior direction $\vec c_{\mathrm{prior}}$
  is the TSR-evolved vacuum state
  $e^{-iH_m}W^\dagger\ket{e_0}$, computed by
  \texttt{tsr\_evolved\_prior(pipeline)}.
\item \texttt{sample\_hmc(posterior, opts)} and
  \texttt{sample\_hmc\_single(posterior, opts)} --- leapfrog-on-the-sphere
  HMC with Metropolis accept/reject. The sampler renormalizes to
  the unit sphere after every leapfrog step (so the trajectory is
  constrained to the natural manifold of the wavefunctions), and
  is initialized at the prior direction (with a small random
  orthogonal perturbation) to ensure fast burn-in. Uses a
  deterministic splitmix64 + Box--Muller PRNG (no external
  dependencies).
\item \texttt{reconstruct(pipeline, c\_sample)} --- the Phase 5
  tomographic reconstruction (convenience wrapper around
  \texttt{QfmPipeline::decode}).
\end{itemize}

\subsubsection*{Kernel surface wiring (rev 17)}

The rev 16 module was research code; rev 17 wires it into the
production kernel surface:

\begin{itemize}
\item \texttt{prob\_kernel::Session::bayesian\_update(observations, hmc\_opts)
  $\to$ BayesianUpdateReport} --- a session-level op that runs the
  full QMF.tex \S8 algorithm end-to-end (likelihood construction
  + HMC + reconstruction) and returns the HMC diagnostics
  (\texttt{log\_posterior}, \texttt{mean\_likelihood}) plus the
  Phase 5 reconstructed full-resolution image.
\item \texttt{unfer\_protocol::BayesianUpdateRequest /
  BayesianUpdateResult / HmcOptsSpec} --- the serde layer for
  the FFI request and result.
\item \texttt{unfer\_ffi::uk\_bayesian\_update} --- the FFI
  symbol: parses the request JSON, calls the session, drains the
  \texttt{BayesianUpdateResult} via \texttt{uk\_get\_result},
  emits a \texttt{conditioned} event for \texttt{uk\_poll}
  subscribers. Returns 0 on success; \texttt{UK-1001} for
  malformed JSON, \texttt{UK-1004} for an invalid model handle,
  \texttt{UK-5000} for non-QFM models.
\item \texttt{unfer/bayes\_update\_module/} --- a fifth end-to-end
  Austral demo (alongside \texttt{demo\_module} +
  \texttt{qfm\_module} + \texttt{qfm\_tomo\_module} +
  \texttt{data\_source}). JIT-creates a \texttt{qfm\_tomography}
  model (4-pt tetrahedron in $d=4$ with $\texttt{krylov\_dim}=4$
  matching $K_2=4$ for the rev 17 SIRK-whitened basis), runs a
  single Bayesian update on a single observation, drains the
  result. The standard \texttt{UK-4001} negative test verifies
  that revoking either \texttt{uk\_get\_result} or
  \texttt{uk\_bayesian\_update} denies the call.
\item \texttt{bayes-update-e2e} job in
  \texttt{.github/workflows/ci.yml} (mirrors the existing
  \texttt{qfm-tomo-e2e} job).
\end{itemize}

\subsubsection*{SIRK-whitened Krylov basis (rev 17, P6 G)}

The Krylov basis $W$ used by the encode and the Bayesian-update
likelihood operators is now the genuine SIRK-generated
$\texttt{w\_whiten}$ matrix (restricted to the $K_2$
single-excitation Fock rows, with per-row renormalization to keep
the Born-rule likelihood well-defined), instead of the
$K_2 \times \mathrm{rank}$ identity sub-block of rev 14. This is
the lossless round-trip promised by \S\ref{sec:tomographic}: each
column of $W$ is a real TSR-derived linear combination of the
$K_2$ single-excitation modes, not a single mode in isolation.
Constraint documented: $\texttt{krylov\_dim} \ge K_2$ (the SIRK
sequence has $\texttt{krylov\_dim}+1$ rows, so the $K_2$-row
restriction is well-defined only when
$\texttt{krylov\_dim} \ge K_2$).

\subsubsection*{Tests + benchmarks (rev 17)}

The implementation now has \textbf{+10 new tests} (170 $\to$ 180
workspace):

\begin{itemize}
\item \texttt{+2} new pipeline tests for the P6 G SIRK-whitened
  basis: \texttt{w\_basis\_is\_sirk\_whitened\_not\_identity} and
  \texttt{w\_basis\_columns\_are\_unit\_norm}.
\item \texttt{+4} new \texttt{prob\_kernel::session::tests}:
  \texttt{bayesian\_update\_smoke\_qfm\_model},
  \texttt{bayesian\_update\_zero\_observations\_returns\_prior},
  \texttt{bayesian\_update\_non\_qfm\_returns\_internal},
  \texttt{bayesian\_update\_dim\_mismatch\_returns\_qfm\_error}.
\item \texttt{+4} new \texttt{unfer\_ffi/tests/ffi.rs} FFI
  integration tests:
  \texttt{bayesian\_update\_via\_ffi},
  \texttt{bayesian\_update\_via\_ffi\_zero\_observations},
  \texttt{bayesian\_update\_via\_ffi\_on\_non\_qfm\_returns\_5000},
  \texttt{bayesian\_update\_via\_ffi\_bad\_obs\_dim\_returns\_1001}.
\end{itemize}

The \texttt{rev 16 bayesian\_update\_tsr\_recovers\_training\_mode}
test was updated in rev 17 to use the new P6 G acceptance
criterion (HMC sample's overlap with the observation's krylov
state, not $\arg\max = 0$ of the decoded image --- which is no
longer the right invariant under the mixed-basis $W$). A fourth
benchmark group \texttt{bayes\_update\_vs\_n} ($N = 1, 4, 16$)
confirms the $\mathcal{O}(N \cdot m^2)$ per-step HMC cost.

\subsection{Current state and next steps (rev 17, 2026-06-30)}

The algorithm in this paper is \textbf{fully closed} in the
\texttt{unfer} kernel (rev 17, \texttt{commit e3c2479}). Every
section has a working implementation: the Mehler-prior diagonal
\emph{workhorse} \S\ref{sec:impl} is the
\texttt{qfm\_mehler} builtin; the off-diagonal
$\Hgen_{\mathrm{off}}$ coupling is the
\texttt{qfm\_mehler\_offdiag} builtin; the Tomographic Subspace
Recovery pipeline of \S\ref{sec:tomographic} is the
\texttt{qfm} crate + the \texttt{qfm\_tomo\_module} Austral
demo + the \texttt{uk\_evolve} FFI symbol; the Quantum Bayesian
Update of \S\ref{sec:bayes} is the \texttt{qfm::bayes} module +
the \texttt{Session::bayesian\_update} method + the
\texttt{uk\_bayesian\_update} FFI symbol + the
\texttt{bayes\_update\_module} Austral demo. The 5 end-to-end
module demos (\texttt{demo\_module} +
\texttt{qfm\_module} + \texttt{qfm\_tomo\_module} +
\texttt{data\_source} + \texttt{bayes\_update\_module}) all
type-check against the \texttt{UnferKernel} bindings and pass
their positive + UK-4001 negative gates in CI.

\textbf{Test counts (CPU, full sweep 2026-06-30, rev 17):}
unfer workspace \textbf{180} (19 fock\_sirk + 26
nested\_fock\_algebra + 37 prob\_kernel + 30 unfer\_protocol +
33 unfer\_ffi + 35 qfm); 5 module demos (positive + UK-4001
negative); 4 criterion benchmark groups in
\texttt{qfm/benches/pipeline.rs}. Clippy clean, fmt clean. The
GPU path is smoke-tested locally on 14 CUDA tests but is not yet
in CI.

\subsubsection*{Next steps to improve}

The next rounds of work are \emph{not} new algorithm features
(the algorithm is closed). They are
\textbf{operationalization, validation, scaling, and
documentation} of the existing implementation. The recommended
order is:

\begin{enumerate}
\item \textbf{P7 (Productionization, $\sim$1 week).} README
  refresh; CUDA-on-CI smoke; real-data validation of QFM-TSR on
  a canonical dataset (MNIST 8$\times$8, Fashion-MNIST, or 2D
  Ising-model snapshots); runtime check for the rev 17 P6 G
  \texttt{krylov\_dim} $\ge K_2$ constraint (currently
  doc-only); \texttt{HmcOptsSpec} FFI validation; Bayesian-update
  benchmark expansion (vary $m$, $K_2$, leapfrog steps in
  addition to $N$).

\item \textbf{P8 (User-facing features, $\sim$2--3 weeks).}
  Typst-math $\to$ Hamiltonian compiler (closes the v1
  documented extension point); \texttt{uk\_belief\_propagation}
  FFI symbol (QMF.tex \S11 message-passing alternative to HMC);
  \texttt{bayes\_update\_module} v2 (multi-observation loop +
  Karcher-mean posterior estimate); CUDA kernel benchmark; a
  new \texttt{iterated\_bayes\_module} Austral demo that drives
  \texttt{uk\_bayesian\_update} + \texttt{uk\_evolve} in a loop.

\item \textbf{P9 (Production infra, $\sim$3--4 weeks).} CUDA
  toolkit pinning (closes the P6 D8 gap); CI: private-repo PAT
  (closes P6 D9); velysterm \texttt{mathed\_mini} Step 4 (caret
  blink, mouse hit-testing, AccessKit bridge, kernel-bridge
  wiring so the Bevy-free frontend can show \texttt{\textbackslash prob}
  results); port the stale \texttt{velyst} examples to retire
  the \texttt{upstream-stale-examples} feature gate.

\item \textbf{P10 (Validation \& science, ongoing).} Scale
  QFM-TSR + Bayesian-update to CIFAR-10 / ImageNet / CFD /
  Ising-model datasets; full Lindbladian
  $e^{(H - i \Sigma L_k^\dagger L_k / 2)\,t}$ solver (currently
  out of scope; the seed exists as the Hermitian off-diagonal
  QFM coupling); a Millennium-Prize mass-gap demonstration on
  the 4D SU(2) Yang-Mills lattice at $l = 3$ (extends the
  $l=2$ P6 A1 test).
\end{enumerate}

The detailed breakdown of every item, the per-item leverage and
effort estimate, and the out-of-scope record are in
\texttt{docs/IMPLEMENTATION\_PLAN.md} \S"Next steps to improve".

\section{Conclusion}

We have given a self-contained derivation of a classical, grid-free realization
of the \emph{wavefunction flow} of Layden, Sweke, Havl\'i\v{c}ek, Chowdhury, and
Neklyudov~\cite{Layden2025} inside the \texttt{unfer} kernel. The paper's
central object --- the continuity Hamiltonian
$\Hgen_t=\tfrac12[\hat p\cdot v_t(\hat x)+v_t(\hat x)\cdot\hat p]$, with the
conservative form $\Hgen_t^c=i[\Kop,V_t(\hat x)]$, fully specified by an
already-learned classical velocity field and generating a qsample
$\Psi_t=\sqrt{p_t}$ --- maps directly onto \texttt{unfer}'s Born-rule Fock
machinery. By encoding data as orthogonal Fock states we replace the paper's
$N^d$ Fourier-collocation grid with an analytical, diagonal,
$\mathcal{O}(M)$ closed-form fit of the conservative potential, with zero data
loss. By compressing the resulting bounded generator with the Shift-Invert
Rational Krylov method --- in place of the paper's product formulas --- we
obtain exponentially convergent, $\mathcal{O}(m^2)$ single-step inference. The
measure-theoretic core --- the Mehler/L\'opez--Temme limit identifying the
uniform hyperspherical measure with the Gaussian measure and Gegenbauer with
Hermite polynomials --- explains \emph{why} the Gaussian source $p_0$ of a flow
model is exactly the Fock vacuum, and why orthogonal excitations carry
information. The shipped builtin realizes the spectrally-stable diagonal
skeleton of this generator; the Hermitian off-diagonal coupling
$\Hgen_{\mathrm{off}}$ of Section~\ref{sec:impl} realizes coherent amplitude
transport between vacuum and data channels; the full \emph{Tomographic
Subspace Recovery} pipeline of Section~\ref{sec:tomographic} decouples
semantics from reasoning via two-level hashing and pre-projected observables,
enabling M-free four-phase online generation at high raw resolution; and the
\emph{Quantum Bayesian Updating} extension of Section~\ref{sec:bayes} shows how
the TSR-evolved prior $P_{\mathrm{prior}}$ can be conditioned on $N$ new
problem-defining observations in $\mathcal{O}(N\,m^2)$ per HMC step, with
no $\mathcal{O}(M)$ dependence, no local Gaussian or mean-field
approximations, and a navigable posterior landscape that would otherwise be
a pure-memorization ``golf course'' under a direct full-Bayesian
integration. The result aligns with the \texttt{unfer} philosophy: rigorous
probability tracking through mathematically exact, geometry-preserving
operations on
separable Hilbert spaces.

\vspace{1em}
\begin{thebibliography}{9}

\bibitem{Layden2025}
Layden, D., Sweke, R., Havl\'i\v{c}ek, V., Chowdhury, A., \& Neklyudov, K.
(2025). \textit{Wavefunction Flows: Efficient Quantum Simulation of Continuous
Flow Models}. arXiv:2510.08462 [quant-ph].

\bibitem{Hashimoto2019}
Hashimoto, Y., \& Nodera, T. (2019). \textit{Shift-invert Rational Krylov
method for an operator $\phi$-function of an unbounded linear operator}.
Japan Journal of Industrial and Applied Mathematics, 36, 1--13.

\bibitem{LopezTemme1999}
L\'opez, J. L., \& Temme, N. M. (1999). \textit{Approximations of orthogonal
polynomials in terms of Hermite polynomials}. Methods and Applications of
Analysis, 6(2), 131--146.

\bibitem{Mehler1866}
Mehler, F. G. (1866). \textit{Ueber die Entwicklung einer Function von
beliebig vielen Variablen nach Laplaceschen Functionen h\"oherer Ordnung}.
Journal f\"ur die reine und angewandte Mathematik, 66, 161--176.

\bibitem{Kopperman2026}
\textit{Beyond Real Numbers: A Hilbert-Space Exchange Format for the UniFormal
(UPL) Framework}. (2026). UniFormal architecture tutorial
(\texttt{Kopperman\textunderscore Tutorial}); Mehler uniform-measure /
Gaussian-limit foundation.

\bibitem{unfer2026}
\texttt{unfer} Kernel Architecture Documentation.
\textit{IMPLEMENTATION\textunderscore PLAN.md, ARCHITECTURE.md, PROTOCOL.md}.
(2026).

\end{thebibliography}

\end{document}
